Towards Practical First-Order Model Counting

### 2024-06-30 doc/paper.tex
New version of FastWFOMC
Notes for improvement

\author{%
  Ananth K. Kidambi$^1$\footnote{The first and second authors contributed equally and were affiliated with the National University of Singapore during the completion of this work.}\and
Guramrit Singh$^1$\and
Paulius Dilkas$^2$\and
Kuldeep S. Meel$^3$ \\
\affiliations
$^1$Indian Institute of Technology Bombay, Mumbai, India\\
$^2$National University of Singapore, Singapore, Singapore\\
$^3$University of Toronto, Toronto, Canada\\
\emails
210051002@iitb.ac.in,
guramrit@iitb.ac.in,
paulius.dilkas@nus.edu.sg,
meel@cs.toronto.edu
}

@@

\section{Introduction}

@@

There are various WFOMC algorithms with different underlying principles. Perhaps
the most prominent class of WFOMC algorithms is based on \emph{first-order
  knowledge compilation}. In this approach, the formula is compiled into a
representation (such as a circuit or graph) by iteratively applying
\emph{compilation rules}. This representation can then be used to compute the
WFOMC for any combination of domain sizes (or weights). Algorithms in this class
include \textsc{ForcLift}~\cite{DBLP:conf/ijcai/BroeckTMDR11} and its recent
extension \textsc{Crane}~\cite{DBLP:conf/kr/DilkasB23}. The former compiles
formulas into circuits, while the latter compiles them first to
\emph{first-order computational graphs} (FCGs) and then to (algebraic)
equations. Another WFOMC algorithm,
\textsc{FastWFOMC}~\cite{DBLP:conf/uai/BremenK21}, is based on cell enumeration.
Other algorithms utilise local search~\cite{DBLP:journals/pvldb/NiuRDS11},
junction trees~\cite{DBLP:conf/aaai/VenugopalSG15}, Monte Carlo
sampling~\cite{DBLP:journals/cacm/GogateD16}, and anytime approximation via
upper/lower bound construction~\cite{DBLP:conf/ijcai/BremenK20}.

@@

The recently proposed \textsc{Crane} algorithm, while capable of handling
formulas beyond the reach of \textsc{ForcLift}, was incomplete as it could only
construct function definitions, which would then need to be evaluated to compute
the WFOMC.\@ Additionally, recursive functions were presented without base
cases. In this work, we introduce \Cranetwo{}, an extension of \textsc{Crane}
that addresses these two weaknesses.

@@

In this section, we will review the basic concepts of first-order logic as they
are used in first-order knowledge compilation algorithms. There are two key
differences between the logic used within such algorithms and the logic
supported as input. First, Skolemization~\cite{DBLP:conf/kr/BroeckMD14} is
employed to eliminate existential quantifiers by introducing additional
predicates. Second, the input formula is rewritten as a conjunction of clauses,
each of which is in \emph{prenex normal form}~\cite{hinman2018fundamentals}.

A \emph{formula} is a conjunction of clauses. A \emph{clause} is of the form
$\forall x_{1} \in \Delta_{1}\text{.}\forall x_{2} \in \Delta_{2}\dots\text{
}\forall x_{n} \in \Delta_{n}\text{.}\phi(x_{1}, x_{2}, \dots, x_{n})$, where
$\phi$ is a disjunction of literals that only contain variables
$x_{1}, \dots, x_{n}$ (and any constants). We say that a clause is a
\emph{(positive) unit clause} if
\begin{enumerate*}[label=(\roman*)]
  \item there is only one literal with a predicate, and
  \item it is a positive literal.
\end{enumerate*}
A \emph{literal} is either an atom (i.e., a \emph{positive} literal) or its
negation (i.e., a \emph{negative} literal). An \emph{atom} is either
\begin{enumerate*}[label=(\roman*)]
  \item $P(t_{1}, \dots, t_{m})$ for some predicate $P$ and terms
  $t_{1}, \dots, t_{m}$ (written as $P(\mathbf{t})$ for short) or
  \item $x=y$ for some terms $x$ and $y$.
\end{enumerate*}
An atom is \emph{ground} if it contains no variables, i.e., only constants. The
\emph{arity} of a predicate is the number of arguments it takes, i.e., $m$ in
the case of predicate $P$ mentioned above. When we want to denote a predicate
together with its arity, we write $P/m$. A \emph{term} is either a variable or a
constant. Throughout the paper, we will use set-theoretic notation, interpreting
a formula as a set of clauses and a clause as a set of literals. Moreover, for
readability, clauses written on separate lines are implicitly conjoined.

@@

@@

In \cref{tbl:logics}, we outline the differences among three first-order logics
commonly used in WFOMC:
\begin{enumerate*}[label=(\roman*)]
  \item \FO{} is the input format for
  \textsc{ForcLift}\footnote{\url{https://github.com/UCLA-StarAI/Forclift}} and
  its extensions
  \textsc{Crane}\footnote{\url{https://doi.org/10.5281/zenodo.8004077}} and
  \Cranetwo{};
  \item \Ctwo{} is often used in the literature on \textsc{FastWFOMC} and
  related
  methods~\cite{DBLP:journals/jair/Kuzelka21,DBLP:conf/aaai/MalhotraS22};
  \item \UFO{} is the input format supported by a private version of
  \textsc{FastWFOMC} obtained directly from the authors.
\end{enumerate*}
Note that the publicly available
version\footnote{\url{https://comp.nus.edu.sg/~tvanbr/software/fastwfomc.tar.gz}}
of \textsc{FastWFOMC} does not support any cardinality constraints. The notation
we use to refer to each logic is standard in the case of \Ctwo{}, new in the
case of \UFO{}, and redefined to be more specific in the case of \FO{}. All
three logics are function-free, and domains are always assumed to be finite.

In \FO{}, each term is assigned to a \emph{sort}, and each predicate $P/n$ is
assigned to a sequence of $n$ sorts. Each sort has its corresponding domain.
Most of these assignments to sorts are typically left implicit and can be
reconstructed from the quantifiers. For example, $\forall x,y \in \Delta\text{.
}P(x, y)$ implies that variables $x$ and $y$ have the same sort. On the other
hand, $\forall x \in \Delta\text{. }\forall y \in \Gamma\text{. } P(x, y)$
implies that $x$ and $y$ have different sorts, and it would be improper to
write, for example, $\forall x \in \Delta\text{. }\forall y \in \Gamma\text{.
} P(x, y) \lor x = y$. \FO{} is also the only logic to support constants,
formulas with more than two variables, and the equality predicate.

@@

\begin{definition}[Model]\label{def:model}
  Let $\phi$ be a formula in \FO{}. For each predicate $P/n$ in $\phi$, let
  ${(\Delta_{i}^{P})}_{i=1}^{n}$ be a list of the corresponding domains (which
  may not be distinct). Let $\sigma$ be a map from the domains of $\phi$ to
  their interpretations as sets, satisfying the following conditions:
  \begin{enumerate*}[label=(\roman*)]
    \item the sets are pairwise disjoint, and
    \item the constants in $\phi$ are included in the corresponding domains.
  \end{enumerate*}
  Then a \emph{structure} of $\phi$ (with respect to $\sigma$) is a set $M$ of
  ground literals defined by adding either $P(\mathbf{t})$ or
  $\neg P(\mathbf{t})$ for every predicate $P/n$ in $\phi$ and $n$-tuple
  $\mathbf{t} \in \prod_{i=1}^{n} \sigma(\Delta_{i}^{P})$. A structure is a
  \emph{model} if it satisfies $\phi$.
\end{definition}

\begin{definition}[WFOMC~\cite{DBLP:conf/ijcai/BroeckTMDR11}]
  Continuing from \cref{def:model}, for each predicate $P$ in $\phi$, let
  $w^{+}(P)$ and $w^{-}(P)$ be its \emph{weights} in $\mathbb{Q}$. Unless
  specified otherwise, we assume all weights are equal to one. The
  \emph{(symmetric) weighted first-order model count} (WFOMC) of $\phi$ (with
  respect to $\sigma$, $w^{+}$, and $w^{-}$) is:
  \[
    \sum_{M \models \phi} \prod_{P(\mathbf{t}) \in M} w^{+}(P) \prod_{\neg P(\mathbf{t}) \in M} w^{-}(P),
  \]
  where the sum is over all models of $\phi$.
\end{definition}

@@

  Equivalently, in \FO{} we would write
  \begin{equation}\label[formula]{eq:fo}
    \begin{gathered}
      \forall x \in \Gamma\text{. }\exists y \in \Delta\text{. }P(x, y)\\
      \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }P(x, y) \land P(x, z) \Rightarrow y = z.
    \end{gathered}
  \end{equation}
  The first clause asserts that each $x$ must have at least one corresponding
  $y$, while the second statement adds the condition that if $x$ is mapped to
  both $y$ and $z$, then $y$ must equal $z$. It is important to note that
  \cref{eq:fo} is written with two domains instead of just one. However, we can
  still determine the number of endofunctions by assuming that the sizes of
  $\Gamma$ and $\Delta$ are equal. This formulation, as observed by Dilkas and
  Belle~\shortcite{DBLP:conf/kr/DilkasB23}, can prove beneficial in enabling
  first-order knowledge compilation algorithms to find efficient solutions.
\end{example}

@@

\section{Completing the Definitions of Recursive Functions}\label{sec:main}

\begin{algorithm}[t]
  \caption{\protect\CompileWithBaseCases{$\phi$}}\label{alg:compilewithbasecases}
  \KwIn{formula $\phi$}
  \KwOut{set $\mathcal{E}$ of equations}
  $(\mathcal{E}, \mathcal{F}, \mathcal{D}) \gets \Compile{$\phi$}$\;\label{line:first}
  $\mathcal{E} \gets \Simplify{$\mathcal{E}$}$\;\label{line:second}
  \ForEach{base case $f(\mathbf{x}) \in \FindBaseCases{$\mathcal{E}$}$}{
    $\psi \gets \mathcal{F}(f)$\;
    \ForEach{$i$ such that $x_{i} \in \mathbb{N}_{0}$}{
      $\psi \gets \Propagate{$\psi$, $\mathcal{D}(f, i)$, $x_i$}$\;
    }
    $\mathcal{E} \gets \mathcal{E} \cup \CompileWithBaseCases{$\psi$}$\;
  }
\end{algorithm}

@@

\begin{example}[Counting bijections]\label{example:overall}
  Consider the following formula (previously examined by Dilkas and
  Belle~\shortcite{DBLP:conf/kr/DilkasB23}) that defines predicate $P$ as a
  bijection between two sets $\Gamma$ and $\Delta$:
  \[
    \begin{gathered}
      \forall x \in \Gamma\text{. }\exists y \in \Delta\text{. }P(x, y)\\
      \forall y \in \Delta\text{. }\exists x \in \Gamma\text{. }P(x, y)\\
      \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }P(x, y) \land P(x, z) \Rightarrow y = z\\
      \forall x, z \in \Gamma\text{. }\forall y \in \Delta\text{. }P(x, y) \land P(z, y) \Rightarrow x = z.
    \end{gathered}
  \]
  We specifically examine the first solution returned by \Cranetwo{} for this
  formula.

@@

Finally, we highlight the crucial role of the \Simplify procedure in simplifying
the algebraic pattern $\sum_{m=0}^{n}[a \le m \le b] f(m)$. Here:
\begin{enumerate*}[label=(\roman*)]
  \item $n$ is a variable,
  \item $a, b \in \mathbb{N}_{0}$ are constants,
  \item $f$ is an expression that may depend on $m$, and
  \item $[a \le m \le b] =
  \begin{cases}
    1 & \text{if $a \le m \le b$} \\
    0 & \text{otherwise}
  \end{cases}$ represents the Iverson bracket.
\end{enumerate*}
\Simplify transforms this pattern into
$f(a) + f(a+1) + \cdots + f(\min\{\, n, b \,\})$. For instance, in the case of
\cref{example:overall}, \Simplify transforms
$g(l, m) = \sum_{k=0}^{m}[0 \le k \le 1]\binom{m}{k}g(l-1, m-k)$ into the
simplified form mentioned above.

@@

\begin{theorem}[Termination]\label{thm:halting}
  Let $f$ be an $n$-ary function in $\mathcal{E}$ and
  $\mathbf{x} \in \mathbf{N}_{0}^{n}$. Then the evaluation of $f(\mathbf{x})$
  terminates.
\end{theorem}

@@

\subsection{Smoothing for Base Cases}\label{sec:smoothingbase}

In this section, we motivate and describe \cref{line:smoothing1,line:smoothing2}
of \cref{alg:propagate}. Suppose that \Propagate is called with arguments
$(\phi, \Delta, 0)$, which means we are simplifying the formula $\phi$ by
assuming that the domain $\Delta$ is empty. Informally, if there is a predicate
$P$ in $\phi$ that has nothing to do with the domain $\Delta$, smoothing
preserves all occurrences of $P$ even if all clauses with $P$ become vacuously
satisfied. It is important to note that the approach presented in this section
is not unique. We explain it via the example below.

@@

\section{Experimental Evaluation}\label{sec:experiments}

@@

\begin{figure}[t]
  \centering
  \input{plot}
  \caption{The running time of WFOMC algorithms. Note that the $y$-axis is on a
    logarithmic scale.}\label{fig:plot}
\end{figure}

@@

As shown in \cref{fig:plot}, the runtimes of all compilation-based algorithms
remained practically constant, in contrast to the rapidly increasing runtimes of
\textsc{FastWFOMC}. Although the search/compilation part is slower in
\Cranetwo{} than in \textsc{ForcLift}, the difference is negligible. The
runtimes of the knowledge compilation algorithms appear constant because, for
these counting problems and domain sizes, compilation time dominates inference
time (remember that compilation time is independent of domain sizes). Indeed,
the maximum inference time of \Cranebfs{} and \Cranegreedy{} across these
experiments is only \SI{4}{\milli\second}. The runtimes of \Cranetwo{} have
lower variation than those of \textsc{ForcLift} because we compile the formula
anew for each domain size with \textsc{ForcLift}, whereas with \Cranetwo{}, we
compile it once and reuse the resulting C++ program for all domain sizes.

>>>>

\author{%
  Ananth K. Kidambi$^1$\footnote{The first and second authors contributed equally and were affiliated with the National University of Singapore during the completion of this work.}\and
Guramrit Singh$^1$\and
Paulius Dilkas$^2$\and
Kuldeep S. Meel$^3$ \\
\affiliations
$^1$Indian Institute of Technology Bombay, Mumbai, India\\
$^2$National University of Singapore, Singapore, Singapore\\
$^3$University of Toronto, Toronto, Canada\\
\emails
210051002@iitb.ac.in,
guramrit@iitb.ac.in,
paulius.dilkas@nus.edu.sg,
meel@cs.toronto.edu
}% TODO: update affiliation and email

@@

% TODO: No proper runtime consideration (other than the termination). (address
% this criticism in some way)

% TODO: Having an algorithm that can efficiently handle large domain sizes is
% crucial in the weighted setting. For example, consider the well-known "friends
% and smokers" Markov logic network [1]. The utility of such a model would be
% significantly limited if exact probabilities could only be efficiently
% computed for social networks of at most 30 people.
% [1] Richardson, M., & Domingos, P. (2006). Markov logic networks. Machine learning, 62, 107-136.

% TODO: somehow indicate that source code will be available (a redacted link?)

% TODO: Updates to the algorithm
% 1) replace x_i-c_i on Line 7 of Algorithm 2 with a generalisation of x_i that can be an arbitrary summation/subtraction of domain sizes
% 2) a function call with a constant should always mean that that function should have a base case, not just when it's in the same function

\section{Introduction}
% TODO: improvements to the introduction (must):
% * make it extra clear what the gap is that this work is trying to fill
% * mention Crane earlier and describe its importance
% * the last two paragraphs should make more of a BOOM (e.g., include the over 9k claim)
% * (here and elsewhere) clarify the distinction between the weighted and unweighted variants of the problem in the final version of the paper. We focus on the unweighted variant in our implementation and experiments, however, Crane2 can handle WFOMC instances as well simply by switching the main data type from integers to rational numbers. Maybe add some weighted experiments as well.

@@

There are various WFOMC algorithms with different underlying principles. Perhaps
the most prominent class of WFOMC algorithms is based on \emph{first-order
  knowledge compilation}. In this approach, the formula is compiled into a
representation (such as a circuit or graph) by iteratively applying
\emph{compilation rules}. This representation can then be used to compute the
WFOMC for any combination of domain sizes (or weights). Algorithms in this class
include \textsc{ForcLift}~\cite{DBLP:conf/ijcai/BroeckTMDR11} and its recent
extension \textsc{Crane}~\cite{DBLP:conf/kr/DilkasB23}. The former compiles
formulas into circuits, while the latter compiles them first to
\emph{first-order computational graphs} (FCGs) and then to (algebraic)
equations. Another WFOMC algorithm,
\textsc{FastWFOMC}~\cite{DBLP:conf/uai/BremenK21}, is based on cell enumeration.
Other algorithms utilise local search~\cite{DBLP:journals/pvldb/NiuRDS11},
junction trees~\cite{DBLP:conf/aaai/VenugopalSG15}, Monte Carlo
sampling~\cite{DBLP:journals/cacm/GogateD16}, and anytime approximation via
upper/lower bound construction~\cite{DBLP:conf/ijcai/BremenK20}.
% TODO: FCGs were never defined. Maybe avoid referring to them altogether?
% Smoothing nodes are also undefined.

@@

% TODO: first sentence: elaborate a bit more for a general audience
The recently proposed \textsc{Crane} algorithm, while capable of handling
formulas beyond the reach of \textsc{ForcLift}, was incomplete as it could only
construct function definitions, which would then need to be evaluated to compute
the WFOMC.\@ Additionally, recursive functions were presented without base
cases. In this work, we introduce \Cranetwo{}, an extension of \textsc{Crane}
that addresses these two weaknesses.
% TODO: We called the previous version of Crane "incomplete" because it cannot
% be used as a black box that takes Markov logic networks (or weighted formulas
% in some other format) and outputs numbers (i.e., model counts or
% probabilities).

@@

In this section, we will review the basic concepts of first-order logic as they
are used in first-order knowledge compilation algorithms. There are two key
differences between the logic used within such algorithms and the logic
supported as input. First, Skolemization~\cite{DBLP:conf/kr/BroeckMD14} is
employed to eliminate existential quantifiers by introducing additional
predicates. Second, the input formula is rewritten as a conjunction of clauses,
each of which is in \emph{prenex normal form}~\cite{hinman2018fundamentals}.
% TODO: explain that "Skolemization" here is NOT standard Skolemization of
% first-order logic, which would replace existential quantifiers by function
% terms.

% TODO: reorder the definitions in the opposite direction
A \emph{formula} is a conjunction of clauses. A \emph{clause} is of the form
$\forall x_{1} \in \Delta_{1}\text{.}\forall x_{2} \in \Delta_{2}\dots\text{
}\forall x_{n} \in \Delta_{n}\text{.}\phi(x_{1}, x_{2}, \dots, x_{n})$, where
$\phi$ is a disjunction of literals that only contain variables
$x_{1}, \dots, x_{n}$ (and any constants). We say that a clause is a
\emph{(positive) unit clause} if
\begin{enumerate*}[label=(\roman*)]
  \item there is only one literal with a predicate, and
  \item it is a positive literal.
\end{enumerate*}
A \emph{literal} is either an atom (i.e., a \emph{positive} literal) or its
negation (i.e., a \emph{negative} literal). An \emph{atom} is either
\begin{enumerate*}[label=(\roman*)]
  \item $P(t_{1}, \dots, t_{m})$ for some predicate $P$ and terms
  $t_{1}, \dots, t_{m}$ (written as $P(\mathbf{t})$ for short) or
  \item $x=y$ for some terms $x$ and $y$.
\end{enumerate*}
An atom is \emph{ground} if it contains no variables, i.e., only constants. The
\emph{arity} of a predicate is the number of arguments it takes, i.e., $m$ in
the case of predicate $P$ mentioned above. When we want to denote a predicate
together with its arity, we write $P/m$. A \emph{term} is either a variable or a
constant. Throughout the paper, we will use set-theoretic notation, interpreting
a formula as a set of clauses and a clause as a set of literals. Moreover, for
readability, clauses written on separate lines are implicitly conjoined.

@@

% TODO: mention the Unique Name Assumption and that, hence, s = t holds iff s
% and t are the same constant.

@@

% TODO:
% 1. Rename \UFO{} to what that new paper uses
% 2. Use FastWFOMC.jl instead, removing the mention of 'privately-obtained'
% 3. use symbols instead of numbers for the footnotes
In \cref{tbl:logics}, we outline the differences among three first-order logics
commonly used in WFOMC:
\begin{enumerate*}[label=(\roman*)]
  \item \FO{} is the input format for
  \textsc{ForcLift}\footnote{\url{https://github.com/UCLA-StarAI/Forclift}} and
  its extensions
  \textsc{Crane}\footnote{\url{https://doi.org/10.5281/zenodo.8004077}} and
  \Cranetwo{};
  \item \Ctwo{} is often used in the literature on \textsc{FastWFOMC} and
  related
  methods~\cite{DBLP:journals/jair/Kuzelka21,DBLP:conf/aaai/MalhotraS22};
  \item \UFO{} is the input format supported by a private version of
  \textsc{FastWFOMC} obtained directly from the authors.
\end{enumerate*}
Note that the publicly available
version\footnote{\url{https://comp.nus.edu.sg/~tvanbr/software/fastwfomc.tar.gz}}
of \textsc{FastWFOMC} does not support any cardinality constraints. The notation
we use to refer to each logic is standard in the case of \Ctwo{}, new in the
case of \UFO{}, and redefined to be more specific in the case of \FO{}. All
three logics are function-free, and domains are always assumed to be finite.

In \FO{}, each term is assigned to a \emph{sort}, and each predicate $P/n$ is
assigned to a sequence of $n$ sorts. Each sort has its corresponding domain.
Most of these assignments to sorts are typically left implicit and can be
reconstructed from the quantifiers. For example, $\forall x,y \in \Delta\text{.
}P(x, y)$ implies that variables $x$ and $y$ have the same sort. On the other
hand, $\forall x \in \Delta\text{. }\forall y \in \Gamma\text{. } P(x, y)$
implies that $x$ and $y$ have different sorts, and it would be improper to
write, for example, $\forall x \in \Delta\text{. }\forall y \in \Gamma\text{.
} P(x, y) \lor x = y$. \FO{} is also the only logic to support constants,
formulas with more than two variables, and the equality predicate.
% TODO: the discussion about sorts isn't featured in the rest of the paper

@@

% TODO: The distinctness of domains is important in two ways. First, in terms of
% expressiveness, writing clause such as "\forall x, P(x, x)" is valid if
% predicate P is defined over two copies of the same domain and invalid
% otherwise. Second, having more distinct domains makes the problem more
% decomposable for the knowledge compilation algorithm. With more distinct
% domains, it can make assumptions or deductions about, e.g., the first domain
% of predicate P without worrying how (or if) they apply to the second domain.
\begin{definition}[Model]\label{def:model}
  Let $\phi$ be a formula in \FO{}. For each predicate $P/n$ in $\phi$, let
  ${(\Delta_{i}^{P})}_{i=1}^{n}$ be a list of the corresponding domains (which
  may not be distinct). Let $\sigma$ be a map from the domains of $\phi$ to
  their interpretations as sets, satisfying the following conditions:
  \begin{enumerate*}[label=(\roman*)]
    \item the sets are pairwise disjoint, and
    \item the constants in $\phi$ are included in the corresponding domains.
  \end{enumerate*}
  Then a \emph{structure} of $\phi$ (with respect to $\sigma$) is a set $M$ of
  ground literals defined by adding either $P(\mathbf{t})$ or
  $\neg P(\mathbf{t})$ for every predicate $P/n$ in $\phi$ and $n$-tuple
  $\mathbf{t} \in \prod_{i=1}^{n} \sigma(\Delta_{i}^{P})$. A structure is a
  \emph{model} if it satisfies $\phi$.
\end{definition}
% TODO: adding to what

% TODO: “Weights in Q” undefined, C2 undefined (and also far from well explained
% before), endofunction undefined (also endomorphisms missing), weight function
% undefined, weighted model count imprecise (above it’s called weighted
% first-order model count/wfmoc),
\begin{definition}[WFOMC~\cite{DBLP:conf/ijcai/BroeckTMDR11}]
  Continuing from \cref{def:model}, for each predicate $P$ in $\phi$, let
  $w^{+}(P)$ and $w^{-}(P)$ be its \emph{weights} in $\mathbb{Q}$. Unless
  specified otherwise, we assume all weights are equal to 1. The
  \emph{(symmetric) weighted first-order model count} (WFOMC) of $\phi$ (with
  respect to $\sigma$, $w^{+}$, and $w^{-}$) is:
  \[
    \sum_{M \models \phi} \prod_{P(\mathbf{t}) \in M} w^{+}(P) \prod_{\neg P(\mathbf{t}) \in M} w^{-}(P),
  \]
  where the sum is over all models of $\phi$.
\end{definition}

@@

  % TODO: Yes, as mentioned at the end of Section 2.1, clauses c and d written
  % on separate lines are to be interpreted as the conjunction of c and d (maybe
  % move that sentence here?).
  Equivalently, in \FO{} we would write
  \begin{equation}\label[formula]{eq:fo}
    \begin{gathered}
      \forall x \in \Gamma\text{. }\exists y \in \Delta\text{. }P(x, y)\\
      \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }P(x, y) \land P(x, z) \Rightarrow y = z.
    \end{gathered}
  \end{equation}
  The first clause asserts that each $x$ must have at least one corresponding
  $y$, while the second statement adds the condition that if $x$ is mapped to
  both $y$ and $z$, then $y$ must equal $z$. It is important to note that
  \cref{eq:fo} is written with two domains instead of just one. However, we can
  still determine the number of endofunctions by assuming that the sizes of
  $\Gamma$ and $\Delta$ are equal. This formulation, as observed by Dilkas and
  Belle~\shortcite{DBLP:conf/kr/DilkasB23}, can prove beneficial in enabling
  first-order knowledge compilation algorithms to find efficient solutions.
\end{example}

@@

\section{Completing the Definitions of Recursive Functions}\label{sec:main}
% TODO: As mentioned in the caption of Figure 1, "Compile" refers to the
% knowledge compilation algorithm of Dilkas and Belle (2023). Its three return
% values are explained at the beginning of page 4. We do not elaborate on the
% details of "Simplify" as it primarily consists of elementary algebraic
% simplifications such as "1x = x" and "0x = 0". However, we highlight one
% non-trivial aspect of "Simplify" right before Section 3.1. We will improve the
% structure of Section 3 to make this information more readily available. Maybe
% write Crane instead of Compile?

% TODO: It would not be entirely accurate to call the algorithms presented in
% this work a "post-processing step" because Algorithm 1 is recursive. In other
% words, this "post-processing step" to first-order knowledge compilation can
% lead to more compilation that then updates/completes the original set of
% equations.
\begin{algorithm}[t]
  \caption{\protect\CompileWithBaseCases{$\phi$}}\label{alg:compilewithbasecases}
  \KwIn{formula $\phi$}
  \KwOut{set $\mathcal{E}$ of equations}
  $(\mathcal{E}, \mathcal{F}, \mathcal{D}) \gets \Compile{$\phi$}$\;\label{line:first}
  $\mathcal{E} \gets \Simplify{$\mathcal{E}$}$\;\label{line:second}
  \ForEach{base case $f(\mathbf{x}) \in \FindBaseCases{$\mathcal{E}$}$}{
    $\psi \gets \mathcal{F}(f)$\;
    \ForEach{$i$ such that $x_{i} \in \mathbb{N}_{0}$}{
      $\psi \gets \Propagate{$\psi$, $\mathcal{D}(f, i)$, $x_i$}$\;
    }
    $\mathcal{E} \gets \mathcal{E} \cup \CompileWithBaseCases{$\psi$}$\;
  }
\end{algorithm}
% TODO: >> Alg 1, line 5: Why is this not an infinite loop? On line 5 of
% Algorithm 1, the loop iterates over the indices of \mathbf{x}. The arity of a
% function is always a finite number, typically equal to 1 or 2.

@@

% TODO: F missing (also not explained that it’s not needed below). E is not explained other than that we take the example from X. Would not be a problem if it is motivated differently and the focus is on post-processing (defining inputs properly) and not on the entire algorithm.
% TODO: include a high-level overview of first-order knowledge compilation and the meaning of the resulting functions. For the most part, the compilation process that constructs these function definitions can be treated as a black box for the purposes of this work. However, a few high-level details still ought to be mentioned:
% 1. Each algebraic variable ("m", "n", and "l" in the example) corresponds to a domain size.
% 2. \mathcal{D} indicates that "m" denotes the size of \Gamma, "n" denotes the size of \Delta, etc.
% 3. \mathcal{E} can contain any (positive) number of functions.
% 4. Function denoted by "f" represents the solution to the model counting problem.
% 3. The model count of the original formula is computed by evaluating f on the desired domain sizes. For example, f(3, 2) is the model count when |\Gamma| = 3, and |\Delta| = 2.
\begin{example}[Counting bijections]\label{example:overall}
  Consider the following formula (previously examined by Dilkas and
  Belle~\shortcite{DBLP:conf/kr/DilkasB23}) that defines predicate $P$ as a
  bijection between two sets $\Gamma$ and $\Delta$:
  \[
    \begin{gathered}
      \forall x \in \Gamma\text{. }\exists y \in \Delta\text{. }P(x, y)\\
      \forall y \in \Delta\text{. }\exists x \in \Gamma\text{. }P(x, y)\\
      \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }P(x, y) \land P(x, z) \Rightarrow y = z\\
      \forall x, z \in \Gamma\text{. }\forall y \in \Delta\text{. }P(x, y) \land P(z, y) \Rightarrow x = z.
    \end{gathered}
  \]
  We specifically examine the first solution returned by \Cranetwo{} for this
  formula.

@@

% TODO: maybe omit the word 'iverson'...
Finally, we highlight the crucial role of the \Simplify procedure in simplifying
the algebraic pattern $\sum_{m=0}^{n}[a \le m \le b] f(m)$. Here:
\begin{enumerate*}[label=(\roman*)]
  \item $n$ is a variable,
  \item $a, b \in \mathbb{N}_{0}$ are constants,
  \item $f$ is an expression that may depend on $m$, and
  \item $[a \le m \le b] =
  \begin{cases}
    1 & \text{if $a \le m \le b$} \\
    0 & \text{otherwise}
  \end{cases}$ represents the Iverson bracket.
\end{enumerate*}
\Simplify transforms this pattern into
$f(a) + f(a+1) + \cdots + f(\min\{\, n, b \,\})$. For instance, in the case of
\cref{example:overall}, \Simplify transforms
$g(l, m) = \sum_{k=0}^{m}[0 \le k \le 1]\binom{m}{k}g(l-1, m-k)$ into the
simplified form mentioned above.

@@

\begin{theorem}[Termination]\label{thm:halting}
  Let $f$ be an $n$-ary function in $\mathcal{E}$ and
  $\mathbf{x} \in \mathbb{N}_{0}^{n}$. Then the evaluation of $f(\mathbf{x})$
  terminates.
\end{theorem}

@@

\subsection{Smoothing for Base Cases}\label{sec:smoothingbase}
% TODO: To be precise, line 7 of Algorithm 3 checks whether smoothing is
% necessary, and lines 8-9 execute it. If the condition on line 7 is not
% satisfied, the clause is not smoothed but rather completely omitted.

In this section, we motivate and describe \cref{line:smoothing1,line:smoothing2}
of \cref{alg:propagate}. Suppose that \Propagate is called with arguments
$(\phi, \Delta, 0)$, which means we are simplifying the formula $\phi$ by
assuming that the domain $\Delta$ is empty. Informally, if there is a predicate
$P$ in $\phi$ that has nothing to do with the domain $\Delta$, smoothing
preserves all occurrences of $P$ even if all clauses with $P$ become vacuously
satisfied. It is important to note that the approach presented in this section
is not unique. We explain it via the example below.
% TODO: In this context, "not unique" means that there are many different ways
% to adjust/rewrite the formula to achieve the goal of smoothing. This section
% (and Algorithm 3) presents one such approach.

@@

\section{Experimental Evaluation}\label{sec:experiments}
% TODO: A WFOMC instance comprises of a formula in first-order logic, a pair of
% weights for each predicate, and a domain size (i.e., a non-negative integer)
% for each domain. We provide the full details of all of these, including how
% the formulas ought to be rewritten to fit the format of each WFOMC algorithm.

% TODO: must rerun on an HPC

% TODO: "hard to follow", pointing to the DPMC paper as a good example

% TODO: Compilation time is included in the experimental results, and the
% differences between compilation and inference time are delineated in the
% second-to-last paragraph of Section 6.

% TODO: It would not have been feasible to increase domain sizes by 1 all the
% way up to 300000. Thus, the experimental results are presented in two parts,
% where the first part examines scalability with small domain sizes, and the
% second part tests the limits of each algorithm within a given timeout.

% TODO: maybe instead of a table I should have an experiment where:
% 1) Crane scales exponentially (with a base of sqrt(2))
% 2) other algorithm scale linearly
% 3) Then I can just draw the usual cumulative plot
% 4) Stop each algorithm when they reach the timeout value

% TODO: (here or in the introduction) C++ code generation is necessary for any
% runtime experiments because without it the algorithm is "incomplete" as
% explained above.

% TODO: I'm claiming that C++ code generation is not the main contribution, but
% the experiments are mainly about it...

% TODO: The asymptotic analysis of the total number of arithmetic operations
% required to compute the FOMC of these formulas was already performed in
% previous work [1].

% TODO: Since FastWFOMC does not support many-sorted logic, our experiments were restricted to formulas with a single domain. However, many of the counting problems examined in [1] become equivalent (in the sense that they generate the same integer sequences) when restricted to a single domain. Moreover, it is challenging to compare Crane2 and FastWFOMC on a larger set of problems because each problem has to be represented in two (fairly different) logics: FO and UFO^2+EQ.

@@

\begin{figure}[t]
  \centering
  \input{plot}
  \caption{The running time of WFOMC algorithms. Note that the $y$-axis is on a
    logarithmic scale.}\label{fig:plot}
\end{figure}
% TODO: turn this into two separate plots for clarity. Maybe also attach labels
% to their corresponding lines

@@

As shown in \cref{fig:plot}, the runtimes of all compilation-based algorithms
remained practically constant, in contrast to the rapidly increasing runtimes of
\textsc{FastWFOMC}. Although the search/compilation part is slower in
\Cranetwo{} than in \textsc{ForcLift}, the difference is negligible. The
runtimes of the knowledge compilation algorithms appear constant because, for
these counting problems and domain sizes, compilation time dominates inference
time (remember that compilation time is independent of domain sizes). Indeed,
the maximum inference time of \Cranebfs{} and \Cranegreedy{} across these
experiments is only \SI{4}{\milli\second}. The runtimes of \Cranetwo{} have
lower variation than those of \textsc{ForcLift} because we compile the formula
anew for each domain size with \textsc{ForcLift}, whereas with \Cranetwo{}, we
compile it once and reuse the resulting C++ program for all domain sizes.
% TODO: The lines for Crane2 in Figure 4 appear to be completely flat whereas
% the line for ForcLift has some slight fluctuations. We refer to these
% fluctuations as variance.
